---
title: "Exercise 3.1.1 Basic Data Wrangling"
author: "James Hooi"
date: "October 25, 2017"
output: html_document
---

<style type="text/css">
body{ font-size: 12px; }
td { font-size: 8px; }
h1.title { font-size: 38px; color: DarkBlue; }
h1 { font-size: 28px; }
h2 { font-size: 20px; }
h3 { font-size: 16px; }
code.r{ font-size: 11px; line-height: 13px; }
pre { font-size: 11px; line-height: 13px; }
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=500)

# Load required libraries
library(dplyr)
library(tidyr)
```

## Before wrangling
Load the data and review how it looks.

```{r dataload, eval=TRUE, echo=TRUE, warning=FALSE}
  orig.csv <- read.csv("refine_original.csv", stringsAsFactors = FALSE)
  print(orig.csv)
```

## Wrangling steps

Using dplyr and tidyr, clean the data per instructions.

- 1: Clean up brand names

Look up name patterns and create cleaned_company column with the cleaned names.

```{r wrangling01, eval=TRUE, echo=TRUE, warning=FALSE}
  company.distct <- orig.csv %>% distinct(tolower(trimws(company))) %>% select(company = contains("company"))
  company.distct <- company.distct %>%
    mutate(cleaned_company = case_when(
                                      grepl(pattern = "ips$", .$company) | grepl(pattern = "^phi", .$company) ~ "philips",
                                      grepl(pattern = "^ak", .$company) ~ "akzo",
                                      grepl(pattern = "^van", .$company) ~ "van houten",
                                      grepl(pattern = "^uni", .$company) ~ "unilever"
                                    ))
  print(company.distct)
```

Join cleaned names back to main data frame, select columns, clean column names. Then:

- 2: Separate product code and number
- 3: Add product categories (p = Smartphone, v = TV, x = Laptop, q = Tablet)
- 4: Add full address for geocoding
- 5: Create dummy variables for company and product category

```{r wrangling02, eval=TRUE, echo=TRUE, warning=FALSE}
  working.df <- inner_join(orig.csv, company.distct, by="company") %>%
    select(company = cleaned_company, product_code_number = Product.code...number, address:name ) %>%
    # 2: Separate product code and number ####
    separate(., product_code_number, into = c("product_code","product_number"), sep="-") %>%
    # 3: Add product categories (p = Smartphone, v = TV, x = Laptop, q = Tablet) ####
    mutate( product_category = case_when(
                                        product_code == "p" ~ "Smartphone",
                                        product_code == "v" ~ "TV",
                                        product_code == "x" ~ "Laptop",
                                        product_code == "q" ~ "Tablet"
                                    )
    # 4: Add full address for geocoding ####
            , full_address = paste(address, city, country, sep=", ")
    # 5: Create dummy variables for company and product category ####
            , company_philips = if_else(company=="philips",1,0)
            , company_akzo = if_else(company=="akzo",1,0)
            , company_van_houten = if_else(company=="van houten",1,0)
            , company_unilever = if_else(company=="unilever",1,0)
            , product_smartphone = if_else(product_category=="Smartphone",1,0)
            , product_tv = if_else(product_category=="TV",1,0)
            , product_laptop = if_else(product_category=="Laptop",1,0)
            , product_tablet = if_else(product_category=="Tablet",1,0)
          )
```

## After wrangling

Cleaned data (3 outputs):
```{r outputdata1, eval=TRUE, echo=TRUE, warning=FALSE}
  print(working.df[,1:8])
  print(working.df[,9:13])
  print(working.df[,14:17])
```
