---
title: "Exercise 3.1.2 Dealing with Missing Values"
author: "James Hooi"
date: "October 25, 2017"
output: html_document
---

<style type="text/css">
body{ font-size: 12px; }
td { font-size: 8px; }
h1.title { font-size: 38px; color: DarkBlue; }
h1 { font-size: 28px; }
h2 { font-size: 20px; }
h3 { font-size: 16px; }
code.r{ font-size: 11px; line-height: 13px; }
pre { font-size: 11px; line-height: 13px; }
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=500)

# Load required libraries
library(dplyr)
library(tidyr)
```

## Before wrangling
Load the data and review how it looks. Fix missing values per instructions.

```{r dataload, eval=TRUE, echo=TRUE, warning=FALSE}
  orig.csv <- read.csv("titanic_original.csv", stringsAsFactors = FALSE)
  str(orig.csv)
```

### 1: Port of embarkation is missing for passengers from Southampton

```{r explore1, eval=TRUE, echo=TRUE, warning=FALSE}
rbind(orig.csv %>% group_by(embarked) %>% tally() %>% ungroup, orig.csv[is.na(orig.csv$embarked),])
```

### 2: Age is missing for some passengers (will be replaced with mean)

```{r explore2, eval=TRUE, echo=TRUE, warning=FALSE}
cbind(sum(is.na(orig.csv$age)), sum(!is.na(orig.csv$age)))
```

### 3: Lifeboat is missing for many passengers (will be replaced with dummy value "None")

```{r explore3, eval=TRUE, echo=TRUE, warning=FALSE}
cbind(sum(orig.csv$boat=="" | is.na(orig.csv$boat)), nrow(orig.csv)-sum(orig.csv$boat=="" | is.na(orig.csv$boat)))
```

### 4. Cabin number is missing for many passengers (will be turned into a categorical variable)

```{r explore4, eval=TRUE, echo=TRUE, warning=FALSE}
cbind(sum(is.na(orig.csv$cabin) | orig.csv$cabin==""), nrow(orig.csv)-sum(is.na(orig.csv$cabin) | orig.csv$cabin==""))
```

## After wrangling

The issues identified above were fixed per the exercise instructions.

```{r fix, eval=TRUE, echo=TRUE, warning=FALSE}

working.df <- orig.csv %>%
  # 1: Port of embarkation - replace missing values with Southampton
  mutate(embarked = if_else(embarked=="","S", embarked)
  # 2: Age - replace missing values with the mean
         , age = if_else(is.na(age), mean(age, na.rm=TRUE), age)
  # 3: Lifeboat - replace missing with dummy value
         , boat = if_else(boat=="", "None", boat)
  # 4: Cabin - create new column has_cabin_number: 1 if there is a cabin number, 0 otherwise
         , has_cabin_number = if_else(cabin=="", 0, 1 ))

head(working.df[,1:8], 50)
head(working.df[,9:15], 50)
```

